---
title: "506EDAandModeling"
author: "Andrew Zazueta"
date: "12/1/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Setting Work Directory, Loading Libraries, and Loading Data

```{r, warning=FALSE, message=FALSE}
setwd("C:/Users/mzazu/OneDrive/Documents/USD papers/506/AirQualityUCI")
library('astsa')
library('xlsx')
library('tidyverse')
library('lubridate')
library('imputeTS')
library('Rcpp')
aqClean <- read_csv('AirQualityCleaned.csv')
```

# Exploritory Data Analysis

Now, lets look at some of the time series plots.

### Time Series Plots

```{r}
par(mfrow = c(3,1))
tsplot(x = aqClean$DateTime, y = aqClean$CO.GT., 
       main = "True hourly averaged concentration CO in mg/m^3", ylab = "", xaxt="none", 
       cex.main = 0.85, cex.axis = 0.85, cex.lab = 0.85)
tsplot(x = aqClean$DateTime, y = aqClean$C6H6.GT., 
       main = "True hourly averaged Benzene concentration in microg/m^3", ylab = "", 
       xaxt="none", cex.main = 0.85, cex.axis = 0.85, cex.lab = 0.85)
tsplot(x = aqClean$DateTime, y = aqClean$NOx.GT., 
       main = "True hourly averaged NOx concentration in ppb", ylab = "", xaxt="none", 
       cex.main = 0.85, cex.axis = 0.85, cex.lab = 0.85)
```

After taking a look at a few time series in the data set, it is apparent that it would be better to look at daily average than to look at hourly average. Doing this would make the data more clear and less noisy, and also allow for better forecasting. 

```{r, warning=FALSE}
aqClean$Day <- format(aqClean$DateTime, format = "%Y-%m-%d")
aqDaily <- aggregate(aqClean, list(as.Date(aqClean$Day)), FUN=mean) 
aqDaily <- aqDaily[-c(14, 15)]
names(aqDaily)[1] <- "Day"
```

```{r}
# Now lets take a look at the same graphs now
par(mfrow = c(3,1))
tsplot(x = aqDaily$Day, y = aqDaily$CO.GT., 
       main = "True daily averaged concentration CO in mg/m^3", ylab = "", xlab = "Time (days)",
       cex.main = 0.85, cex.axis = 0.85, cex.lab = 0.85, xaxt="n")
tsplot(x = aqDaily$Day, y = aqDaily$C6H6.GT., 
       main = "True daily averaged Benzene concentration in microg/m^3", ylab = "", 
       cex.main = 0.85, cex.axis = 0.85, cex.lab = 0.85, xlab = "Time (days)", xaxt="none")
tsplot(x = aqDaily$Day, y = aqDaily$NOx.GT., 
       main = "True daily averaged NOx concentration in ppb", ylab = "", xlab = "Time (days)",  
       cex.main = 0.85, cex.axis = 0.85, cex.lab = 0.85, xaxt="none")
```

These graphs are much easier to read, so now we can move on to making the time series stationary.

### Differencing Data

```{r}
# Same graphs for reference
par(mfrow = c(3,1))
tsplot(diff(aqDaily$CO.GT.), 
       main = "True daily growth concentration of CO in mg/m^3", ylab = "", xlab = "Time (days)",
       cex.main = 0.85, cex.axis = 0.85, cex.lab = 0.85, xaxt="n")
tsplot(diff(aqDaily$C6H6.GT.), 
       main = "True daily growth of Benzene concentration in microg/m^3", ylab = "", 
       cex.main = 0.85, cex.axis = 0.85, cex.lab = 0.85, xlab = "Time (days)", xaxt="none")
tsplot(diff(aqDaily$NOx.GT.), 
       main = "True daily growth of NOx concentration in ppb", ylab = "", xlab = "Time (days)",  
       cex.main = 0.85, cex.axis = 0.85, cex.lab = 0.85, xaxt="none")
```

The differencing appeared to make our data stationary, so now we can move on to building an ARIMA model from our data. We can also create a cross correlation graph (CCF) between different features to determine if they have similar traits. For example, seeing the CCF between temperature and other features can yield interesting results as to what atmospheric chemicals contribute to temperature changes. 

### Looking at P/ACF and CCF for Different Features

```{r}
# CO in mg/m^3
acf2(diff(diff(aqDaily$CO.GT., 7))) # Weekly fluctuations
```

Seasonal: The ACF is cutting off a lag 7s (s = 7), whereas the PACF is tailing off at lags 7s, 14s, 21s. These results imply an SMA(7), P = 0, Q = 7.

Non-Seasonal: It appears that the ACF cuts off at lag 7, whereas the PACF tails off. This suggests an MA(7) with p = 0 and q = 7.

```{r}
# Benzene in microg/m^3 
acf2(diff(aqDaily$C6H6.GT., 7)) # Appears to have a weekly spike
```

Seasonal: The PACF is cutting off a lag 1s (s = 7), whereas the ACF is tailing off at lags 1s, 7s, 14s. These results imply an SAR(1), P = 1, Q = 0.

Non-Seasonal: It appears that the PACF cuts off at lag 1, whereas the ACF tails off. This suggests an AR(1) with p = 1 and q = 0.

# Building Models and Forecasting

### Determining the ARMA models

```{r}
# For CO
sarima(aqDaily$CO.GT., p = 0, d = 1, q = 7, P = 0, D = 1, Q = 7, S = 7)
```

It appears to be alright. Some of our Q-statistic values are less than 0.05, but the majority are not. 

```{r, warning=FALSE}
# For Benzene
sarima(aqDaily$C6H6.GT., p = 1, d = 1, q = 1, P = 1, D = 0, Q = 1, S = 7)
```

Adding p = 1 and P = 1 to our model improves the Q-statistic values.

### Forecasting 

```{r}
# Plotting the final month of time series for CO
x <- ts(aqDaily$CO.GT., start = decimal_date(as.Date("2004-03-10")), frequency = 365)
CO <- window(x, start=decimal_date(as.Date("2004-03-10")),
            end=decimal_date(as.Date("2005-03-04")))
sarima.for(CO, n.ahead = 32, p = 0, d = 1, q = 7, P = 0, D = 1, Q = 7, S = 7, plot.all=FALSE,
           main = "Forecasting CO")
text(decimal_date(as.Date("2005-02-04")), 5, "PAST")
text(decimal_date(as.Date("2005-03-20")), 5, "FUTURE")
abline(v=decimal_date(as.Date("2005-03-04")), lty=2, col=4)
lines(x)
```

The 'Future' portion of our graph is where the forecasting is made. The black line represents the actual data collected while the red line shows our ARMA model prediction. The grayed out areas are confidence intervals. 

```{r}
# Plotting the final month of time series for Benzene
x <- ts(aqDaily$C6H6.GT., start = decimal_date(as.Date("2004-03-10")), frequency = 365)
C6H6 <- window(x, start=decimal_date(as.Date("2004-03-10")),
            end=decimal_date(as.Date("2005-03-04")))
sarima.for(C6H6, n.ahead = 32, p = 1, d = 1, q = 1, P = 1, D = 0, Q = 1, S = 7,
           plot.all=FALSE, main = "Forecasting Benzene")
text(decimal_date(as.Date("2005-02-04")), 22, "PAST")
text(decimal_date(as.Date("2005-03-20")), 22, "FUTURE")
abline(v=decimal_date(as.Date("2005-03-04")), lty=2, col=4)
lines(x)
```
```{r}
#Plotting the final month of time series for NOx
x <- ts(aqDaily$NOx.GT., start = decimal_date(as.Date("2004-03-10")), frequency = 365)
NOx<- window(x, start=decimal_date(as.Date("2004-03-10")),
            end=decimal_date(as.Date("2005-03-04")))
sarima.for(NOx, n.ahead = 32, p = 1, d = 1, q = 1, P = 1, D = 0, Q = 1, S = 7,
           plot.all=FALSE, main = "Forecasting NOx")
text(decimal_date(as.Date("2005-02-04")), 22, "PAST")
text(decimal_date(as.Date("2005-03-20")), 22, "FUTURE")
abline(v=decimal_date(as.Date("2005-03-04")), lty=2, col=4)
lines(x)
```

```{r}
#Plotting  the final month of time series for Nitrogen Oxide(NO2)
x <- ts(aqDaily$NO2.GT., start = decimal_date(as.Date("2004-03-10")), frequency = 365)
NO2<- window(x, start=decimal_date(as.Date("2004-03-10")),
            end=decimal_date(as.Date("2005-03-04")))
sarima.for(NO2, n.ahead = 32, p = 1, d = 1, q = 1, P = 1, D = 0, Q = 1, S = 7,
           plot.all=FALSE, main = "Forecasting Nitrogen Oxide")
text(decimal_date(as.Date("2005-02-04")), 22, "PAST")
text(decimal_date(as.Date("2005-03-20")), 22, "FUTURE")
abline(v=decimal_date(as.Date("2005-03-04")), lty=2, col=4)
lines(x)
```

```{r}
#Plotting the final month of time series for indium oxide"
x <- ts(aqDaily$PT08.S5.O3., start = decimal_date(as.Date("2004-03-10")), frequency = 365)
PT08.S5<- window(x, start=decimal_date(as.Date("2004-03-10")),
            end=decimal_date(as.Date("2005-03-04")))
sarima.for(PT08.S5 , n.ahead = 32, p = 1, d = 1, q = 1, P = 1, D = 0, Q = 1, S = 7,
           plot.all=FALSE, main = "Forecasting indium oxide")
text(decimal_date(as.Date("2005-02-04")), 22, "PAST")
text(decimal_date(as.Date("2005-03-20")), 22, "FUTURE")
abline(v=decimal_date(as.Date("2005-03-04")), lty=2, col=4)
lines(x)
```

```{r}
#Plotting the final month of time series for tin oxide "
x <- ts(aqDaily$PT08.S1.CO., start = decimal_date(as.Date("2004-03-10")), frequency = 365)
tin_oxide<- window(x, start=decimal_date(as.Date("2004-03-10")),
            end=decimal_date(as.Date("2005-03-04")))
sarima.for(tin_oxide , n.ahead = 32, p = 1, d = 1, q = 1, P = 1, D = 0, Q = 1, S = 7,
           plot.all=FALSE, main = "Forecasting  tin oxide")
text(decimal_date(as.Date("2005-02-04")), 22, "PAST")
text(decimal_date(as.Date("2005-03-20")), 22, "FUTURE")
abline(v=decimal_date(as.Date("2005-03-04")), lty=2, col=4)
lines(x)
```





